---
title: "Reproducible Research Project 1"
author: "Giovanni Madejski"
date: "June 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(ggplot2)
library(scales)
library(Hmisc)
```
# Introduction
 This is the first project for the Reproducible Research course In the Coursera Data Science Specialization. In this project we will be answering a series of questions about a dataset containing data from a fitbit.
 
# Loading and Preprocessing Data
 
 Download, unzip, and load dataset into a dataframe
 
```{r echo = TRUE}
 
 if(!file.exists("repdata%2Fdata%2Factivity.zip")) {
        temp <- tempfile()
        download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
        unlink(temp)
}

data <- read.csv("activity.csv")
```

# What is the mean total number of steps taken per day?

Calculate steps per day, create histogram, find mean and median.

```{r echo=TRUE}
dailysteps <- aggregate(steps ~ date, data, sum)
hist(dailysteps$steps, main = paste("Total Steps Each Day"), col = "Green", xlab="Number of Steps")
stepmean <- mean(dailysteps$steps)
stepmedian <- median(dailysteps$steps) 
```

The mean is `r round(stepmean, digits = 2)`.

The median is `r stepmedian`.


# What is the average daily activity pattern?

Calculate average steps per interval, Plot the average number of steps per day, Find interval with most steps.

```{r echo = TRUE}
intervalsteps <- aggregate(steps ~ interval, data, mean)
plot(intervalsteps$interval, intervalsteps$steps, type = "l", xlab = "Interval", ylab = "Number of Steps", main = "Average Number of Steps per Interval")
maximum <- intervalsteps[which.max(intervalsteps$steps), 1]
```
The 5 minute interval containg the maximum average number of steps across days is `r maximum`.

# Input missing values
Report the number of missing values
```{r echo = TRUE}
missingvalues <- length(which(is.na(data$steps)))
```
The number of missing values is `r missingvalues`

Create a Dataset that fills in missing values with a mean value
```{r echo = TRUE}
filleddata <- data
filleddata$steps <- impute(data$steps, fun=mean)
```

# Create a Histogram of the total number of steps taken each day and Calculate the mean and median steps taken per day

```{r echo = TRUE}
stepsByDay <- tapply(filleddata$steps, filleddata$date, sum)
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency', binwidth = 500)

stepsByDaymean <- mean(stepsByDay)
stepsByDaymedian <- median(stepsByDay)
```
The Mean is `r round(stepsByDaymean,2)` the Median is `r round(stepsByDaymedian,2)`.

Imputting mean data for the missing values has brought the median closer to the mean value. This makes sense as now there are a greater number of "mean" entries in the dataset.

# Are there differences in activty patterns between weekdays and weekends?
Seperate data into weekends and weekdays, then create a panel plot to compare.
```{r echo=TRUE}
filleddata$dateType <-  ifelse(as.POSIXlt(filleddata$date)$wday %in% c(0,6), 'weekend', 'weekday')
averagedfilleddata <- aggregate(steps ~ interval + dateType, data=filleddata, mean)
ggplot(averagedfilleddata, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("interval") + 
    ylab("avarage number of steps")
```

This completes the project, Thank you for reading. 




